\section*{Acknowledgements}    

I would like to thank my supervisor, J.J. Collins for being a fantastic support for not just this project, but throughout the past 2 years, and despite his efforts though, he still hasn't gotten rid of all the commas from my writing style, as this sentence proves! I'd also like to thank Norah Power, for her aid in getting me back into college and her support once I got back in. I can't not mention Chris Exton too, for his fantastic support, advice and guidance in my first semester back after ten years away from academic life. A general thank you is required for the CSIS lecturers in general who are always willing to lend an ear, and point, or push, you in the right direction. 

I owe a debt I can never repay to my parents. Their support through the past number of years has been nothing short of amazing, and I'm very lucky to have them. My parents-in-law, and grandparents-in-law, have also been a fantastic support and the best babysitting service that anyone could ask for! To my class, even though I joined halfway though your four years, you made me feel like I had been there the whole time, and it has been, and will continue to be, a pleasure to know you, especially Shane, who partnered with me on every project, and helped me get to where I am now, thanks!

Last, and certainly not least, my own family, my wife Michelle and son Zach. Zach will be glad to see me finish college so I can get a job and buy him Lego. Almost there!

Michelle, you've been a support beyond words the past few years, and putting up with me these last few months can't have been easy, but you get to see this in print so it's not all bad. 

You were right.